---
output: github_document
bibliography: inst/REFERENCES.bib
link-citations: yes
linkcolor: cyan
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# <img src="man/figures/nashlogo.svg" style='width: 100%; object-fit: contain'/>

<!-- badges: start -->
```{r, include=FALSE}
usethis::use_github_action_check_standard()
badger::badge_license("GPL-3")
badger::badge_code_size("ThomasDelSantoONeill/nash")
```

[![R-CMD-check](https://github.com/ThomasDelSantoONeill/nash/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ThomasDelSantoONeill/nash/actions/workflows/R-CMD-check.yaml)
[![License: GPL-3](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://cran.r-project.org/web/licenses/GPL-3)
[![](https://img.shields.io/github/languages/code-size/ThomasDelSantoONeill/nash.svg)](https://github.com/ThomasDelSantoONeill/nash)
<!-- badges: end -->

# Quick Start

The goal of `nash` is to compute @Nash1951 Equilibrium (NE) harvesting rates for naturally occurring species that **biologically interact** through *e.g.* predation and/or competition. [NE](https://en.wikipedia.org/wiki/Nash_equilibrium) harvesting rates mean *to exploit each species in a wild mixture at such rates that no deviation from this rate can increase the long-term yield from that species* [@Farcas2016].

The algorithms implemented in `nash` assume that **users have developed an ecosystem model** with $S$ harvested compartments, typically represented by population biomass variables. In the **simplest case**, population dynamics are given by a system of autonomous ordinary differential equations of the general form:

$$\frac{d\mathbf{B}}{dt}=\mathbf{f}(\mathbf{B})\circ\mathbf{B}-\mathbf{F}\circ\mathbf{B},$$

with $\mathbf{B}$ representing the non-negative biomass or state vector (dimensions $\text{MASS}$), $\mathbf{f}(\mathbf{B})$ specifying the population growth (or decay) rate in the absence of exploitation (dimensions $1/\text{TIME}$) and the last term describing removal (*e.g.* harvesting and/or culling) at a rate $\mathbf{F}$ (dimensions $1/\text{TIME}$). In addition, $\circ$ denotes the entry-wise or [Hadamard](https://en.wikipedia.org/wiki/Hadamard_product_(matrices)) product.

To run `nash`, the user is required to define an `R` function that runs the above model for given $\mathbf{F}$ values and returns yields at the stable equilibrium $\mathbf{Y}=\mathbf{F}\circ\mathbf{B}^*=\mathbf{F}\circ\mathbf{B}^*(\mathbf{F})$. The `nash` function will then approximate the model near equilibrium via a multispecies Lotka-Volterra (LV) model, for which the NE can be computed analytically and so a first estimation of optimal $\mathbf{F}$ obtained. Subsequently, an updated LV approximation is calculated near the equilibrium given by this new $\mathbf{F}$. `nash` will then re-compute the NE starting a new iteration until a (user-adjustable) convergence threshold for $\mathbf{F}$ is reached.

# Installation

You can install the development version of `nash` either through the `devtools` [@devtools2022] or `remotes` [@remotes2023] packages: 

``` r
# install.packages("devtools")
remotes::install_github("ThomasDelSantoONeill/nash")
```

# Minimal Example

The following chunk of code implements a modified two-species competitive LV model as `HQLV` and showcases the execution of `nash`. For this example, the `HQLV` function also includes a numerical integration solver via the `deSolve` package of @deSolve2010 that returns long-term yields for given harvesting rates.

``` {r}
# Load libraries .
library (deSolve) # ODE solver library
library (nash)
# Initial conditions and parameters .
y <- c(b1 = 0.02, b2 = 0.001)
parameters <- c(r1 = 1, r2 = 1,
                a11 = 1, a12 = 0.5,
                a21 = 0.25, a22 = 1)
time <- 1:100
# Numerical fudge to avoid biomasses becoming negative .
inv <- 1e-5
# Model formulation .
HQLV <- function(par, avg.window = 10) {
  derivs <- function(time, y, parameters) {
    with (as.list(c(y, parameters)), {
      db1.dt = b1*(r1-a11*b1-a12*b2^2) - par[1]*b1 + inv
      db2.dt = b2*(r2-a22*b2-a21*b1^2) - par[2]*b2 + inv
      return(list(c(db1.dt, db2.dt)))
      })
  }
  # Default integrator in deSolve
  simulation <- ode(y = y, times = time, func = derivs,
                    parms = c(parameters, par))
  # Yield computation
  yields <- array(dim = c(nrow(simulation), length(par)))
  for(i in 1:nrow(simulation)) {
    yields[i,] <- simulation[i, -1] * par
  }
  return(colMeans(tail(yields, n = avg.window)))
}
# Execution of nash
NE <- nash(par = c(0.2, 0.3), fn = HQLV, progress = FALSE)
# Results
print(NE)
```

By setting the `yield.curves` argument within the `nash` call equal to `TRUE`, it is possible to compute the yield curves at the NE to verify that, indeed, "*no fleet can attain higher yields by changing their corresponding harvesting rates*".

``` {r, echo = FALSE}
NE <- nash(par = c(0.2, 0.3), fn = HQLV, progress = FALSE, yield.curves = TRUE)
par(mfrow = c(1,2))
plot(NE$YieldEQ[[1]][,1],NE$YieldEQ[[1]][,2], type = "l",
     xlab = "",
     ylab = "")
plot(NE$YieldEQ[[2]][,1],NE$YieldEQ[[2]][,2], type = "l",
     xlab = "",
     ylab = "")
mtext(text = "Yield (MASS/TIME)", side = 2, line = -0.25)
mtext(text = "Harvesting Rate (1/TIME)", side = 1, line = 2, adj = 0)

```
# Assistance

If you encounter a bug, please file an issue with a minimal reproducible example on [GitHub](https://github.com/ThomasDelSantoONeill/nash/issues). For questions and other discussions/enhancements please email me at [t.j.delsantooneill@qmul.ac.uk](t.j.delsantooneill@qmul.ac.uk).

## References

