---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nash

<!-- badges: start -->
```{r, include=FALSE}
usethis::use_github_action_check_standard()
badger::badge_license("GPL-3")
badger::badge_code_size("ThomasDelSantoONeill/nash")
```

[![R-CMD-check](https://github.com/ThomasDelSantoONeill/nash/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ThomasDelSantoONeill/nash/actions/workflows/R-CMD-check.yaml)
[![License: GPL-3](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://cran.r-project.org/web/licenses/GPL-3)
[![](https://img.shields.io/github/languages/code-size/ThomasDelSantoONeill/nash.svg)](https://github.com/ThomasDelSantoONeill/nash)
<!-- badges: end -->

The goal of `nash` is to compute Nash equilibrium (NE)<sup>[[1]](#1)</sup> harvesting rates ($\mathbf{F}$; dimensions $1/\text{TIME}$) for ecological models of the form:

$$
\frac{d\mathbf{B}}{dt}=\mathbf{f}(\mathbf{B})\circ\mathbf{B}-\mathbf{F}\circ\mathbf{B},
$$

with $\mathbf{B}$ representing the non-negative biomass state vector (dimensions $\text{MASS}$), $\mathbf{f}(\mathbf{B})$ specifying the population growth (or decay) rate in the absence of exploitation (dimensions $1/\text{TIME}$) and $\circ$ denoting the entry-wise or [Hadamard](https://en.wikipedia.org/wiki/Hadamard_product_(matrices)) product.

To run `nash`, the user is required to define an `R` function that contains the above model alongside an integration routine to solve it; that, for given $\mathbf{F}$, the user-specified model outputs the yields at the stable equilibrium $\mathbf{Y}=\mathbf{F}\circ\mathbf{B}^*=\mathbf{F}\circ\mathbf{B}^*(\mathbf{F})$. The `nash` function will then approximate the model near equilibrium dynamics by a multispecies Lotka-Volterra (LV) model, for which the NE can be computed analytically and so a first estimation of optimal $\mathbf{F}$ obtained. Subsequently, an updated LV approximation is calculated near the equilibrium given by this new $\mathbf{F}$. `nash` will then re-compute the NE starting a new iteration until a (user-adjustable) convergence threshold for $\mathbf{F}$ is reached.

## Installation

You can install the development version of `nash` either through the `devtools`<sup>[[2]](#2)</sup> or `remotes`<sup>[[3]](#3)</sup> packages: 

``` r
# install.packages("devtools")
devtools::install_github("ThomasDelSantoONeill/nash")
```

## Assistance

If you encounter a clear bug, please file an issue with a minimal reproducible example on [GitHub](https://github.com/ThomasDelSantoONeill/nash/issues). For questions and other discussions/enhancements please email me at [t.j.delsantooneill@qmul.ac.uk](t.j.delsantooneill@qmul.ac.uk).

## References
<a id="1">[1]</a> 
Nash, J. (1951). 
Non-cooperative games. 
<i>Annals of Mathematics</i>, 54(2), 286--295 ([https://doi.org/10.2307/1969529](https://doi.org/10.2307/1969529)).

<a id="2">[2]</a>
Wickham, H., Hester, J. and Chang, W. (2020).
<i>devtools: Tools to Make Developing R Packages Easier</i>.
R package version 2.3.1.
([https://github.com/r-lib/devtools](https://github.com/r-lib/devtools)).

<a id="3">[3]</a>
Hester, J., Csárdi, G., Wickham, H., Chang, W., Morgan, M. and Tenenbaum, D. (2020).
<i>remotes: R Package Installation from Remote Repositories, Including “GitHub”</i>.
R package version 2.2.0.
([https://github.com/r-lib/remotes](https://github.com/r-lib/remotes)).
