---
output: github_document
bibliography: inst/REFERENCES.bib
link-citations: yes
linkcolor: cyan
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# <img src="man/figures/nashlogo.svg" style='width: 100%; object-fit: contain'/>

<!-- badges: start -->
```{r, include=FALSE}
usethis::use_github_action_check_standard()
badger::badge_license("GPL-3")
badger::badge_code_size("ThomasDelSantoONeill/nash")
```

[![R-CMD-check](https://github.com/ThomasDelSantoONeill/nash/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ThomasDelSantoONeill/nash/actions/workflows/R-CMD-check.yaml)
[![License: GPL-3](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://cran.r-project.org/web/licenses/GPL-3)
[![](https://img.shields.io/github/languages/code-size/ThomasDelSantoONeill/nash.svg)](https://github.com/ThomasDelSantoONeill/nash)
<!-- badges: end -->

# Quick Start

The goal of `nash` is to compute @Nash1951 Equilibrium (NE) harvesting rates for naturally occurring species that **biologically interact** through *e.g.* predation and/or competition. [NE](https://en.wikipedia.org/wiki/Nash_equilibrium) harvesting rates mean *to exploit each species in a wild mixture at such rates that no deviation from this rate can increase the long-term yield from that species* [@Farcas2016].

The algorithms implemented in `nash` assume that **users have developed an ecosystem model** with $S$ harvested compartments, typically represented by population biomass variables. In the **simplest case**, population dynamics are given by a system of autonomous ordinary differential equations of the general form:

$$\frac{d\mathbf{B}}{dt}=\mathbf{f}(\mathbf{B})\circ\mathbf{B}-\mathbf{F}\circ\mathbf{B},$$

with $\mathbf{B}$ representing the non-negative biomass or state vector (dimensions $\text{MASS}$), $\mathbf{f}(\mathbf{B})$ specifying the population growth (or decay) rate in the absence of exploitation (dimensions $1/\text{TIME}$) and the last term describing removal (*e.g.* harvesting and/or culling) at a rate $\mathbf{F}$ (dimensions $1/\text{TIME}$). In addition, $\circ$ denotes the entry-wise or [Hadamard](https://en.wikipedia.org/wiki/Hadamard_product_(matrices)) product.

- Add Gustav's comments and aknowledge him.

To run `nash`, the user is required to define an `R` function that contains the above model alongside an integration routine to solve it; that, for given $\mathbf{F}$, the user-specified model outputs the yields at the stable equilibrium $\mathbf{Y}=\mathbf{F}\circ\mathbf{B}^*=\mathbf{F}\circ\mathbf{B}^*(\mathbf{F})$. The `nash` function will then approximate the model near equilibrium dynamics by a multispecies Lotka-Volterra (LV) model, for which the NE can be computed analytically and so a first estimation of optimal $\mathbf{F}$ obtained. Subsequently, an updated LV approximation is calculated near the equilibrium given by this new $\mathbf{F}$. `nash` will then re-compute the NE starting a new iteration until a (user-adjustable) convergence threshold for $\mathbf{F}$ is reached.

## Installation

You can install the development version of `nash` either through the `devtools` [@devtools2022] or `remotes` [@remotes2023] packages: 

``` r
# install.packages("devtools")
devtools::install_github("ThomasDelSantoONeill/nash")
```

## Minimal Example

## Assistance

If you encounter a clear bug, please file an issue with a minimal reproducible example on [GitHub](https://github.com/ThomasDelSantoONeill/nash/issues). For questions and other discussions/enhancements please email me at [t.j.delsantooneill@qmul.ac.uk](t.j.delsantooneill@qmul.ac.uk).

## References

